---
title: "NTOLab14"
author: "NTO"
date: "4/23/2019"
output: html_document
---

```{r,echo = FALSE,include=FALSE}

library(tidyverse)
library(lubridate)
library(stringr)
library(modelr)
options(na.action = na.warn)

babies <- read_csv("https://raw.githubusercontent.com/ervance1/Sp2018-Data-Science-repo/master/babies2a.dat")
babies <- rename(babies, bwtoz = `bwt/oz`)

```

****
### Overall questions and Importance: 
****

* "Do mothers' smoking habits affect the presence of premature births?" and "Do mothers' smoking habits affect the correlation between their babys' gestational age and birth weight?"

* These questions are important because answering them will allow us to provide recommendations to mothers on whether or not smoking is hurtful for their newborn babies.  Most parents care deeply about the well-being of their babies, who contain fragile health, making this statistical study very relevant and important. 

****
### Team Section
****

```{r,echo = FALSE}
mod2 <- lm(`bwt/oz` ~ gestation * smoke, data = davie)

gridtry <- davie %>% 
 data_grid(gestation,smoke) %>% 
  gather_predictions(mod2)

co <- babies%>%na.omit()%>%
  group_by(smoke)%>%
  nest()

smoke_mod <- function(df){
  lm(`bwt/oz`~gestation,data = df)
}

co <- co%>%
  mutate(mod = map(data, smoke_mod))

co$mod[[1]]
co$mod[[2]]

ggplot(davie, aes(gestation,`bwt/oz`)) + 
  geom_point(alpha = 1/4, color = "blue") + facet_grid(~smoke)+
  geom_line(data = gridtry, aes(y = pred),size = 1)+ggtitle("Gestation and Birth Weight of Smokers vs Non-Smokers")+ylab("bwtoz")

```


* Findings and Recommendations: From our subquestions we found that mother's smoking habits do affect the presence of premature births. When comparing gestation and birth weight, babies are predicted to have a smaller birth weight when the mother smokes during pregnancy compared to mother's that never smoke. We recommend not smoking during pregnancy to prevent lower birth weights, and births.

* New Tools / Comparison with Lab 4: 

****
### David's Section
****

Subquestion and Importance: Is there a coorelation between gestation and birthweight? Are ther other underlying facors that effect this coorelation?

This question is important to address because typically a premature babies are charachterized by a low gestation periods and premature babies usually weight less. If there are other factors that effect this coorelation, it can give us insight on what other variables have an effect on prematureness. 

```{r,echo = FALSE}

dave1 <- babies %>% filter(!is.na(gestation))

model1 <- function(a,data){
  a[1]+data* a[2]
    #linear model a +xb where x is data set
  # data is in form data$column
}

distance <- function(a,data){
  diff <- data - model1(a,data)
  abs(mean(diff))
    # data is in form data$column
    # finds distance
}

bwt_dist <- function(a1,a2,data){
  distance(c(a1,a2),dave1$bwtoz)
}
```

```{r,echo= FALSE}
doop <- dave1[,c(4,3)]
#best <- optim(c(0,0),distance, data = doop)
#best$par
## intial starting point in the c(0,0). finds the best slope to the data 

mod3 <- lm(bwtoz ~ gestation, data = doop)
a <- coef(mod3)
a

ggplot(data = dave1, aes(gestation,bwtoz))+
  geom_point(color = "grey15")+
 geom_abline(aes(intercept = a[1], slope = a[2]),color = "blue",size = 1.2)+ ggtitle("Birth Weight Vs. Gestation period") + ylab("Birth Weight(oz)")

#grid <- doop %>% data_grid(bwtoz)

#grid <- grid %>% add_predictions(mod3)

#ggplot(doop, aes(bwtoz)) +
 # geom_point(aes(y=gestation))+
  #geom_line( aes( y = pred),data = grid)

```

I used `lm()` to find the coeffients for the line a best fit using the model ` a[1]+data* a[2]`. The intercept and slope are shown in the tibble above. The graph shows a positive trend between gestation and birthweight. The lower the gestation period the lower the birthweight.

```{r,echo=FALSE,include=FALSE}
#dope <- doop %>% add_residuals(mod3)

#ggplot(data = dope,aes(resid))+
#  geom_freqpoly(binwidth = 1)

davie <- dave1[,c(3,4,18)] %>% filter(!is.na(smoke)) %>% mutate(smoke = str_replace_all(smoke, c("1" = "still smokes", "0" = "never")))
ggplot(davie,aes(gestation,bwtoz))+
  geom_point(aes(color = smoke))
```
****
How does a mother's smoking habits influence this coorelation??

```{r,echo=FALSE}
mod2 <- lm(bwtoz ~ gestation * smoke, data = davie)

gridtry <- davie %>% 
 data_grid(gestation,smoke) %>% 
  gather_predictions(mod2)

ggplot(davie, aes(gestation,bwtoz,colour = smoke)) + 
  geom_point() + 
  geom_line(data = gridtry, aes(y = pred),size = 1)+ggtitle("Colored by smoke habits")
```

I used the model `y ~ x1*x2` or `bwtoz ~ gestation * smoke` to see if the smoking habits along with gestation influence birthweight. I used `gather_predictions()` to plot the lines of best fit for each smoking habit. Smoking habit does have an effect on the coorelation between birthweight and gestation. The line of best fit for mothers who still smoke is significally steeper. This means that mothers who smoke have more likelyhood of a premature baby. Also of premature babies, babies with mothers that smoke weight less.

```{r,echo=FALSE}
residdiff <- davie %>% add_residuals(mod2)

ggplot(data = residdiff,aes(resid,color = smoke))+
  geom_freqpoly(binwidth = 2,size = 1)+ xlab("Residual") + ggtitle("Residual distribution in respect to best linear model")

```

I used `add_residuals()` along with my model to plot the distribution of the redisuals in respect to the best linear model. The residial distributions are centered around zero which shows that the lines of best fit are accurate. The highest residual spike for mothers who still smoke as a higher residual. This makes sense if you consider the increased slope for the previous graph.

****
### Ryan's Section
****

Subquestion and Importance: Age and weight on gestation, given premature. we know that smoking causes premature babies, but what about the mother's age and weight, and how do they compare between smokers and non smokers?

```{r,echo = FALSE}
babies3 <- babies %>% filter(Premature == "1")
#find correlation between premature babies and smoking

ggplot((data = babies3), aes(x = mpregwt, y = gestation)) +
  geom_point()
#effect of pregnancy weight on gestation, given premature
mod <- lm(mpregwt ~ gestation, data = babies3) #fits linear model
coef(mod)

grid1 <- babies3 %>%
  data_grid(mpregwt,gestation)

grid1 <- grid1 %>%
  add_predictions(mod)

ggplot(data = babies3, aes(x = mpregwt)) +
  geom_point(aes(y = gestation)) +
  geom_line(aes(y = pred), data = grid1, colour = "red", size =1)

babies3 <- babies3 %>%
  add_residuals(mod1)
#resids very far from 0 so very likely no correlation
ggplot(data = babies3, aes(x =mpregwt , y =gestation)) +
  geom_ref_line(h=0) +
  geom_point()


babies1 <-completeFun(babies, "time")

#time: 0 = never smoked, 1 = still smokes


#new question: given child premature, how does mother's age influnce gestation for smokers and nonsmokers
 

mod <- lm(mage ~ gestation, data = babies3)
coef(mod) 
?coef
grid2 <- babies3 %>%
  data_grid(gestation, mage)

grid2 <- grid2 %>%
  add_predictions(mod)

ggplot(data = babies3, aes(x =mage)) +
  geom_jitter(aes(y = gestation)) +
  geom_line(aes(y = pred), data = grid2, colour = "red", size =1)
#slight negative correlation(no correlation) between age and premature, for smokers
babies3 <- babies3 %>%
  add_residuals(mod1)


ggplot(data = babies3, aes(x = mage, y = resid)) +
  geom_ref_line(h=0) +
  geom_point()


#time: 0 = never smoked, 1 = still smokes

ggplot(data = babies3, aes(x = mage, y = gestation)) +
  geom_jitter() +
  facet_wrap( ~ time)
```

New Tools: grid, add_residuals, add predictions, coef, lm

Findings / Relation to Team Section: yeah so not much of a correlation for either age or mother's weight.

****
### Ethan's Section
****

Subquestion and Importance: Does a mother's pregancy weight correlate with the baby's birth weight and is this correlation affected by smoking habits? This question is important because I can make suggestions to mothers on how to improve health of babies if a correlation exists between these variables.
```{r,echo = FALSE}

babiese <- babies %>%
  na.omit() %>%
  mutate(smokes = ifelse(smoke == "1", "Smokes", "Doesn't Smoke")) %>%
  select(mpregwt, bwtoz, smokes) 

ggplot(data = babiese, aes(x = mpregwt, y = bwtoz)) +
  geom_point() +
  xlab("Pregnancy Weight") +
  ylab("Birth Weight") +
  ggtitle("Mother's Pregnancy Weight vs. Baby's Birth Weight")

model1 <- lm(bwtoz ~ mpregwt, data = babiese)
coef(model1)

grid <- babiese %>%
  data_grid(mpregwt)

stats <- babiese %>%
  add_residuals(model1) %>%
  add_predictions(model1)
```
```{r,echo=FALSE}
#Predictions

ggplot(data = babiese, mapping = aes(x = mpregwt)) +
  geom_point(mapping = aes(y = bwtoz, color = smokes)) +
  geom_line(mapping = aes(y = pred), data = stats, color = "green", size = 2) +
  geom_smooth(mapping = aes(y = bwtoz, color = smokes), method = "lm", se = FALSE, size = 2) + facet_wrap(~smokes) + 
  xlab("Pregnancy Weight") +
  ylab("Birth Weight") +
  ggtitle("Mother's Pregnancy Weight vs. Baby's Birth Weight by Smoking Habit")
  

#Residuals

ggplot(data = stats, mapping = aes(x = mpregwt, y = resid, color = smokes)) +
  geom_smooth(method = "lm", size = 2, se = FALSE) +
  geom_point() +
  facet_wrap(~smokes) +
  xlab("Pregnancy Weight") +
  ylab("Residual") +
  ggtitle("Residuals for Mother's Pregnancy Weight vs. Baby's Birth Weight")

#Nesting

by_smoking <- babiese %>%
  group_by(smokes) %>%
  nest()

smoke_model <- function(df){
  lm(bwtoz ~ mpregwt, data = df)
}

by_smoking <- by_smoking %>%
  mutate(model = map(data, smoke_model)) %>%
  mutate(pred = map(model, predict)) %>%
  mutate(resid = map(model, residuals))

by_smoking$model[[1]]
by_smoking$model[[2]]

```

New Tools: I used nesting to group by smoking habits and get list columns, models, predictions, and residuals for both habits.  I also used lm(), data_grid(), add_predictions(), and add_residuals() to make my basic models.

Findings / Relation to Team Section: I found that a very weak minor positive correlation exists between pregnancy weight and the baby's weight with an R value of 0.17, and smoking habit appears to have some affect.  I created a scatter plot with just pregnancy weight vs. birth weight and then faceted by smoking habit, creating both an overall model (green line) and comapring it with the models for each facet. There was very little difference between the slope of the lines, though the intercept for women who don't smoke is quite a bit higher, implying we can predict a slightly higher baby's birth weight for those women than women who don't smoke at a given pregnancy weight.  The residuals don't offer much as they are very scattered and show no significant patterns for both smoking habits, implying the weak correlation we knew existed.

****
### Arie's Section
****

Subquestion and Importance:  How does smoking affect the pregnancy weight of moms? This question is  important and interesting because children are the most fragile when they are in the whom. It is crucial for Mom's to be cautious of what they do while carrying since it will have a direct affect to the unborn child.

```{r,echo = FALSE}

babies1 <- babies%>%
  filter(!is.na(mht), !is.na(mpregwt), !is.na(number), !is.na(mrace))%>%
  select(marital, mht, mpregwt, gestation, bwtoz, number, smoke)

c<- cor(x = babies1$mpregwt , y= babies1$mht )


babies1_mod <- lm(mht ~mpregwt, data = babies1)

Arie <- babies1%>%
  data_grid(mht, mpregwt)

Arie <- Arie%>%
  add_predictions(babies1_mod)

ggplot(data = babies1, mapping = aes(x = mpregwt))+geom_point(mapping = aes(y = mht))+facet_grid(~smoke)+
  geom_line(mapping = aes(y = pred), data = Arie, color = "purple", size = .5)+
  ylab("Mom's Average Height")+xlab("Mom's Average Pregnancy Weight")+ggtitle("Average Height and Weight of Mother's that Smoke vs. Non-Smokers", subtitle = "0 = Never Smokes, 1 = Smokes")


```

New Tools: I used `lm()`, `data_grid()`, `add_predictions` to create my linear model ontop of the original data.

Findings / Relation to Team Section: I found that there is a weak positive correlection of .4 between the mothers height and pregnancy weight. The data also shows that smoking has little to no effect on the mothers pregnancy weight.

****
### Anderson's Section
****

Subquestion and Importance: 

```{r,echo = FALSE}

```

New Tools: 

Findings / Relation to Team Section: 

****
#### Who Did What
****

* David -
* Arie - 
* Anderson - 
* Ethan - 
* Ryan - 
